---
title: "CFU data"
author: "Ignacio Romero"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, include=FALSE}
library(readxl)
library(agricolae)
library(tidyr)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(MASS)
library(dplyr)
```


# Analysis of the cfu counted

## Experimental design

Must complete

## Method

### Spore trap

### Counting

# About the data

```{r Load data}
cfu_data <- read_xlsx("raw_files/cfu_data.xlsx")
head(cfu_data, 6)
str(cfu_data)
cfu_data$pot <- as.character(cfu_data$pot)
```

```{r Subset by treatment}
#create a transformation of cfu_g_soil before subsetting to save work
cfu_data$log_cfu <- log2(cfu_data$cfu_g_soil)
cfu_data$sqr_cfu <- sqrt(cfu_data$cfu_g_soil)
cfu_data$inv_cfu <- 1/(cfu_data$cfu_g_soil)
#Subset
cfu_actino <- subset(cfu_data, cfu_data$org== "actinobacteria")
cfu_bact <- subset(cfu_data, cfu_data$org== "bacteria")
cfu_fungi<- subset(cfu_data, cfu_data$org== "fungi")
```


# My research question:

**How has the cfu of each organism changed through the different time points?**

Repeated measure ANOVA with cfu as dependent variable and time and group as independent for each organism independently.   

**How did it change according to each treatment?**



# CFU variation over time

```{r ANOVA cfu actino}
summary(aov(cfu_g_soil ~ time*treatment+ Error(pot), data= cfu_actino))

#Shapiro on untransformed data
cfu_actino %>%
  group_by(time) %>%
  shapiro_test(cfu_g_soil) 
ggqqplot(data=cfu_actino, "cfu_g_soil", facet.by = "time")# Not normal

cfu_actino %>%
  group_by(treatment) %>%
  shapiro_test(cfu_g_soil) 
ggqqplot(data=cfu_actino, "cfu_g_soil", facet.by = "treatment")# Not normal

cfu_actino$cfu_g_soil %>% shapiro.test() # Not normal

hist(cfu_actino$cfu_g_soil)# Not normal


#Shapiro on log transformed data
cfu_actino %>%
  group_by(time) %>%
  shapiro_test(log_cfu) 
ggqqplot(data=cfu_actino, "log_cfu", facet.by = "time")# Not normal

cfu_actino %>%
  group_by(treatment) %>%
  shapiro_test(log_cfu) 
ggqqplot(data=cfu_actino, "log_cfu", facet.by = "treatment")# Not normal

cfu_actino$log_cfu %>% shapiro.test() # Not normal

hist(cfu_actino$log_cfu)# Not normal

#Shapiro on square root transformed data
cfu_actino %>%
  group_by(time) %>%
  shapiro_test(sqr_cfu) 
ggqqplot(data=cfu_actino, "sqr_cfu", facet.by = "time")# Not normal

cfu_actino %>%
  group_by(treatment) %>%
  shapiro_test(sqr_cfu) 
ggqqplot(data=cfu_actino, "sqr_cfu", facet.by = "treatment")# Not normal

cfu_actino$sqr_cfu %>% shapiro.test() # Not normal

hist(cfu_actino$sqr_cfu)# Not normal

#Shapiro on inverse transformed data
cfu_actino %>%
  group_by(time) %>%
  shapiro_test(inv_cfu) 
ggqqplot(data=cfu_actino, "inv_cfu", facet.by = "time")# Not normal



cfu_actino %>%
  group_by(treatment) %>%
  shapiro_test(inv_cfu) 
ggqqplot(data=cfu_actino, "inv_cfu", facet.by = "treatment")# Not normal

cfu_actino$inv_cfu %>% shapiro.test() # Not normal

hist(cfu_actino$inv_cfu)# Not normal


###### BOX -COX transformation

bc <- boxcox(aov(cfu_g_soil ~ time*treatment, data= cfu_actino))
lambda <- bc$x[which.max(bc$y)] 
cfu_actino$bc_cfu <- (cfu_actino$cfu_g_soil ^ lambda - 1) / lambda
cfu_actino %>%  shapiro_test(bc_cfu)
hist(cfu_actino$bc_cfu)
ggqqplot(data= cfu_actino, "bc_cfu", facet.by = "time")

```


```{r ANOVA cfu bact}
summary(aov(cfu_g_soil ~ time*treatment+ Error(pot), data= cfu_bact))

#Shapiro on untransformed data
cfu_bact %>%
  group_by(time) %>%
  shapiro_test(cfu_g_soil) 
ggqqplot(data=cfu_bact, "cfu_g_soil", facet.by = "time")# Not normal

cfu_bact %>%
  group_by(treatment) %>%
  shapiro_test(cfu_g_soil) 
ggqqplot(data=cfu_bact, "cfu_g_soil", facet.by = "treatment")# Not normal

cfu_bact$cfu_g_soil %>% shapiro.test() # Not normal

hist(cfu_bact$cfu_g_soil)# Not normal


#Shapiro on log transformed data
cfu_bact %>%
  group_by(time) %>%
  shapiro_test(log_cfu) 
ggqqplot(data=cfu_bact, "log_cfu", facet.by = "time")# Not normal

cfu_bact %>%
  group_by(treatment) %>%
  shapiro_test(log_cfu) 
ggqqplot(data=cfu_bact, "log_cfu", facet.by = "treatment")# Not normal

cfu_bact$log_cfu %>% shapiro.test() # Not normal

hist(cfu_bact$log_cfu)# Not normal

#Shapiro on square root transformed data
cfu_bact %>%
  group_by(time) %>%
  shapiro_test(sqr_cfu) 
ggqqplot(data=cfu_bact, "sqr_cfu", facet.by = "time")# Not normal

cfu_bact %>%
  group_by(treatment) %>%
  shapiro_test(sqr_cfu) 
ggqqplot(data=cfu_bact, "sqr_cfu", facet.by = "treatment")# Not normal

cfu_bact$sqr_cfu %>% shapiro.test() # Not normal

hist(cfu_bact$sqr_cfu)# Not normal

#Shapiro on inverse transformed data
cfu_bact %>%
  group_by(time) %>%
  shapiro_test(inv_cfu) 
ggqqplot(data=cfu_bact, "inv_cfu", facet.by = "time")# Not normal

cfu_bact %>%
  group_by(treatment) %>%
  shapiro_test(inv_cfu) 
ggqqplot(data=cfu_bact, "inv_cfu", facet.by = "treatment")# Not normal

cfu_bact$inv_cfu %>% shapiro.test() # Not normal

hist(cfu_bact$inv_cfu)# Not normal


###### BOX -COX transformation

bc <- boxcox(aov(cfu_g_soil ~ time*treatment, data= cfu_bact))
lambda <- bc$x[which.max(bc$y)] 
cfu_bact$bc_cfu <- (cfu_bact$cfu_g_soil ^ lambda - 1) / lambda
cfu_bact %>%  shapiro_test(bc_cfu)
hist(cfu_bact$bc_cfu)
ggqqplot(data= cfu_bact, "bc_cfu", facet.by = "time")

```


```{r ANOVA cfu fungi}
summary(aov(cfu_g_soil ~ time*treatment+ Error(pot), data= cfu_fungi))

#Shapiro on untransformed data
cfu_fungi %>%
  group_by(time) %>%
  shapiro_test(cfu_g_soil) 
ggqqplot(data=cfu_fungi, "cfu_g_soil", facet.by = "time")# Not normal

cfu_fungi %>%
  group_by(treatment) %>%
  shapiro_test(cfu_g_soil) 
ggqqplot(data=cfu_fungi, "cfu_g_soil", facet.by = "treatment")# Not normal

cfu_fungi$cfu_g_soil %>% shapiro.test() # Not normal

hist(cfu_fungi$cfu_g_soil)# Not normal


#Shapiro on log transformed data
cfu_fungi %>%
  group_by(time) %>%
  shapiro_test(log_cfu) 
ggqqplot(data=cfu_fungi, "log_cfu", facet.by = "time")# Not normal

cfu_fungi %>%
  group_by(treatment) %>%
  shapiro_test(log_cfu) 
ggqqplot(data=cfu_fungi, "log_cfu", facet.by = "treatment")# Not normal

cfu_fungi$log_cfu %>% shapiro.test() # Not normal

hist(cfu_fungi$log_cfu)# Not normal

#Shapiro on square root transformed data
cfu_fungi %>%
  group_by(time) %>%
  shapiro_test(sqr_cfu) 
ggqqplot(data=cfu_fungi, "sqr_cfu", facet.by = "time")# Not normal

cfu_fungi %>%
  group_by(treatment) %>%
  shapiro_test(sqr_cfu) 
ggqqplot(data=cfu_fungi, "sqr_cfu", facet.by = "treatment")# Not normal

cfu_fungi$sqr_cfu %>% shapiro.test() # Not normal

hist(cfu_fungi$sqr_cfu)# Not normal

#Shapiro on inverse transformed data
cfu_fungi %>%
  group_by(time) %>%
  shapiro_test(inv_cfu) 
ggqqplot(data=cfu_fungi, "inv_cfu", facet.by = "time")# Not normal

cfu_fungi %>%
  group_by(treatment) %>%
  shapiro_test(inv_cfu) 
ggqqplot(data=cfu_fungi, "inv_cfu", facet.by = "treatment")# Not normal

cfu_fungi$inv_cfu %>% shapiro.test() # Not normal

hist(cfu_fungi$inv_cfu)# Not normal


###### BOX -COX transformation

bc <- boxcox(aov(cfu_g_soil ~ time*treatment, data= cfu_fungi))
lambda <- bc$x[which.max(bc$y)] 
cfu_fungi$bc_cfu <- (cfu_fungi$cfu_g_soil ^ lambda - 1) / lambda
cfu_fungi %>%  shapiro_test(bc_cfu)
hist(cfu_fungi$bc_cfu)
ggqqplot(data= cfu_fungi, "bc_cfu", facet.by = "time")
```


The data does not follow a normal distribution no matter which transformation, a non-parametric test will be done instead a parametric ANOVA.


```{r kruskal}
kruskal.actino.treat <- kruskal(cfu_actino$cfu_g_soil, cfu_actino$treatment , alpha= 0.05, p.adj= "bonferroni", group= TRUE)
```

Kruskal Wallis' test does not consider repeated measures (time), but Friedman's test can recognize repeated measures and is non parametric. the only problem is that the input is the mean values, losing the variability of the replicates

```{r Friedmann's test}
 actino<- data.frame("cfu"= cfu_actino$cfu_g_soil, "treatment"= cfu_actino$treatment, "time"= cfu_actino$time)

mean.cfu.actino <- actino %>% group_by(time, treatment) %>% summarise(mean.cfu = mean(cfu))
mean.cfu.data <- cfu_data %>% group_by(time, treatment, org) %>% summarise(mean.cfu = mean(cfu_g_soil))

friedmann.actino.treat <- friedman.test(mean.cfu ~ treatment| time , data = mean.cfu.actino)
friedmann.actino.treat$p.value

```



