---
title: "CFU data"
author: "Ignacio Romero"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, include=FALSE}
library(readxl)
library(agricolae)
library(tidyr)
library(tidyverse)
library(ggpubr)
library(rstatix)
```


# Analysis of the cfu counted

## Experimental design

Must complete

## Method

### Spore trap

### Counting

# About the data

```{r Load data}
cfu_data <- read_xlsx("raw_files/cfu_data.xlsx")
head(cfu_data, 6)
str(cfu_data)
cfu_data$pot <- as.character(cfu_data$pot)
```

```{r Subset by treatment}
cfu_actino <- subset(cfu_data, cfu_data$org== "actinobacteria")
cfu_bact <- subset(cfu_data, cfu_data$org== "bacteria")
cfu_fungi<- subset(cfu_data, cfu_data$org== "fungi")
```


# My research question:

**How has the cfu of each organism changed through the different time points?**

Repeated measure ANOVA with cfu as dependent variable and time and group as independent for each organism independently.   

**How did it change according to each treatment?**



# CFU variation over time

```{r ANOVA cfu actino}

summary(aov(cfu_g_soil ~ time*treatment+ Error(pot), data= cfu_actino))

Anova()

cfu_actino %>%
  group_by(time) %>%
  shapiro_test(cfu_g_soil)

```



