---
title: "Soil respiration. Banana project"
author: "Raquel Correa"
date: "2023-10-24"
output: html_document
---

```{r library, include=FALSE}
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(corrplot)
library(agricolae)
library(emmeans)
library(multcomp)
library(multcompView)
library(data.table)
library(tidyverse)
library(phyloseq)
library(vegan)
library(ape)
library(here)
library(readxl)
library(ggpubr)
library(geomtextpath)
library(knitr)
library(gridExtra)
library(Biostrings)
library(seqinr)
library(FSA)
library(agricolae)
library(MASS)
library(grid)
library(gridExtra)
library(emmeans)
library(multcomp)
library(broom)
library(viridis)
library(ALDEx2)
library(FactoMineR)
library(factoextra)
library(kableExtra)
library(scales)
library(dunn.test)
library(car)
library(corrplot)
library(eulerr)
library(ggforce)
library(ggrepel)
library(rstatix)
library(AICcmodavg)
library(ggthemes)
library(patchwork)
library(egg)
library(cowplot)
library(lme4)
library(lmerTest)
```


```{r banana colours, include=FALSE}
treat.colors <- c("control"= "green4" ,"control.amf"  = "darkseagreen" , "chicken.manure"= "royalblue1", "cow.manure" = "red3","mo.pellet"= "darkorchid1","compost"="chocolate4","control.legume"="darkolivegreen4")

treat.order <- c('control', 'control.amf', 'control.legume', "compost", "chicken.manure", "cow.manure", "mo.pellet") 
treat.labels <- c("Control", "Control\nAMF", "Control\nLegume", "Compost", "Chicken\nmanure", "Cow\nmanure", "Pellet")

```


```{r soil respiration, include=FALSE}
soil_resp <- readxl::read_excel("raw_files/soil_respiration.xlsx")

soil_resp_mean <- soil_resp %>% group_by(treatment,  time) %>% 
  summarise(across(.cols=everything(),.fns=list(mean = mean, sd = sd)),.groups = "drop")


soil_resp_mean
```

### Evolution of soil respiration over time 
One way to measure the microbial community of a soil is respiration. Check how much CO2 is released from it.

A) Basal respiration: The falcon tubes (50 ml), with the original soil samples, were placed in jars with
hermetic closure. The evaluation of respiration (CO2 production) was carried out by titration of the
sodium hydroxide (0.5 M) with hydrochloric acid (0.5 M) once the sample has been incubated for 7 days at 25ºC. Basal respiration provides an indication of the microbial activity supported by the organic matter present in the soil. It reflects the metabolism of microorganisms utilizing the existing organic carbon pool.


B) Induced respiration: Induced respiration was carried out once basal respiration had been evaluated (that is, on the same samples). A solution is added of 20% sterile glucose (0.5 ml/sample). Subsequently, the samples were incubated at 25ºC for
2.5 days and the sodium hydroxide (0.5 M) was titrated with hydrochloric acid (0.5 M).Induced respiration provides insights into the potential microbial activity that can be activated in response to changes in the availability of easily degradable carbon sources. It helps researchers understand how the microbial community responds to specific inputs or disturbances.


The results of basal respiration and induced respiration in the same soil sample can be different. These two measurements capture different aspects of microbial activity and provide complementary information about the soil's response to different conditions.

-Basal Respiration:

Reflects the baseline or inherent microbial activity in the soil under normal, undisturbed conditions.
Represents the ongoing metabolism of microorganisms utilizing the existing organic carbon pool in the soil.
Typically measured without the addition of external carbon sources or specific disturbances.

-Induced Respiration:

Reflects the additional microbial activity that can be stimulated by external factors or disturbances.
Involves the addition of a substrate (e.g., glucose) or other stimuli to induce a response from the microbial community.
Provides insights into the potential for increased microbial activity when additional carbon sources are made available

```{r over the time, echo=FALSE}
basal_plotsd <- soil_resp_mean%>% ggplot(aes(x=time, group= treatment, col= treatment, y=mg_co_basal_mean))+
  geom_line(linewidth=1.1)+
  xlab("Time")+
  ylab("mg C-CO2/kg dry soil")+
  scale_color_manual(values= treat.colors)+
  geom_point(size =2)+
  theme_clean()+
  geom_errorbar(aes(ymin=mg_co_basal_mean-mg_co_basal_sd,   ymax=mg_co_basal_mean+mg_co_basal_sd), size =0.5, width=0.2)+
  labs(title = "Basal respiration over the time.", subtitle = "", caption="Mean and SD")+
  theme(plot.title = element_text(size=12))
basal_plotsd

ggsave(basal_plotsd, file="soil_respiration_result/basal_duringtime.png", width= 1900,height= 1200, units= "px")

induced_plotsd <- soil_resp_mean%>% ggplot(aes(x=time, group= treatment, col= treatment, y=mg_co_induced_mean))+
  geom_line(linewidth=1.1)+
  theme_clean()+
  xlab("Time")+
  ylab("mg C-CO2/kg dry soil")+
  scale_color_manual(values= treat.colors)+
  geom_point(size=2)+
  geom_errorbar(aes(ymin=mg_co_induced_mean-mg_co_induced_sd,   ymax=mg_co_induced_mean+mg_co_induced_sd), width=0.2)+
  labs(title = "Induced respiration over the time.", subtitle = "", caption="Mean and SD")+
  theme(plot.title = element_text(size=12))
induced_plotsd
ggsave(induced_plotsd, file="soil_respiration_result/induced_duringtime.png", width= 1900,height= 1200, units= "px")

basal_induced_overtime <- ggarrange(basal_plotsd + theme(legend.position="bottom"), induced_plotsd+ theme(legend.position="none") , ncol = 2)

ggsave(basal_induced_overtime, file="soil_respiration_result/basal_induced_overtime.png", width= 1900,height= 1200, units= "px")

respiration_inducedplot2 <- soil_resp_mean%>% ggplot(aes(x=time, group= treatment, fill= treatment, y=mg_co_induced_mean))+
  geom_col(aes(fill= treatment), position= "dodge")+
  scale_fill_manual(values= treat.colors)+
  labs(title = "Induced respiration.", subtitle = "", caption="Mean")+
  theme_classic()

respiration_inducedplot2

respiration_basdalplot2 <- soil_resp_mean%>% ggplot(aes(x=time, group= treatment, fill= treatment, y=mg_co_basal_mean))+
  geom_col(aes(fill= treatment), position= "dodge")+
  scale_fill_manual(values= treat.colors)+
  labs(title = "Basal respiration.", subtitle = "", caption="Mean")+
  theme_classic()
respiration_basdalplot2

```

### Differences of respiration between treatments at the start of the experiment (t0)


```{r respiration basal and induced t0, echo=FALSE}

respirationt0 <- subset(soil_resp, soil_resp$time == "t0")
#analized basal
shapiro.test(respirationt0$mg_co_basal)#data basal normal
leveneTest(mg_co_basal~ treatment, data=respirationt0) #equal Variance 

anova_basalt0 <- aov(mg_co_basal ~ treatment + rep, data=respirationt0) #signification different anova two ways
anova_basalt0 <- summary(aov(mg_co_basal ~ treatment + rep, data=respirationt0))

pvalue_anova_basalt0 <- anova_basalt0[[1]][["Pr(>F)"]][1] #se tiene que sacar del summary el p-value

tukey_basalt0 <- HSD.test(aov(mg_co_basal ~ treatment + rep, data=respirationt0), "treatment", alpha = 0.05)$groups %>% rownames_to_column(var= "treatment")

tukey_basalt0

meanbasalt0 <- aggregate(mg_co_basal ~ treatment, data=respirationt0, FUN= mean)

plot_basal_t0 <- respirationt0 %>% ggplot(aes(x= factor(treatment, level= treat.order), y= mg_co_basal, fill=treatment))+
  geom_point(pch= 24, size= 7)+ 
  scale_fill_manual(values= treat.colors)+
  xlab("")+
  ylab("mg C-CO2/kg dry soil")+#basal
  stat_summary(fun=mean, aes(x= treatment, y = mg_co_basal), geom= "crossbar", colour= "black", width= 0.7, data=meanbasalt0)+
  geom_text(data=tukey_basalt0, aes(x=treatment, y = mg_co_basal, label= groups), size= 4, colour="black", fontface = "bold",   nudge_x= 0.2, nudge_y= 0.6)+
  theme(axis.text.x = element_text(angle = 0, vjust = 3))+
  labs(subtitle = "" , caption = "Different letters indicate statistically significant differences between each treatment at T0.\nCrossbar indicate the average of the three samples ")+
  guides(fill=guide_legend(title="Treatment"))+
  theme(plot.title = element_text(size=1))+
  scale_x_discrete(labels= treat.labels)+
  ggtitle("Basal respiration at T0", subtitle= paste("ANOVA (HSDTukey) p-value: ", format(pvalue_anova_basalt0, scientific= TRUE, digits=2 ), sep= ""))+
  theme_classic2()


plot_basal_t0
ggsave(plot_basal_t0, file="soil_respiration_result/dif_basal_t0.png", width= 1900,height= 1200, units= "px")

#analized induced
shapiro.test(respirationt0$mg_co_induced)#data induced no normal
soil_resp %>% group_by(treatment) %>% shapiro_test(mg_co_induced) #all data of each time of induced are normal
leveneTest(mg_co_induced~ treatment, data=respirationt0) #equal variance
anova_inducedt0 <- aov(mg_co_induced ~ treatment +rep, data=respirationt0) #signification different 
anova_inducedt0 <- summary(aov(mg_co_induced ~ treatment +rep, data=respirationt0))


pvalue_anova_inducedt0 <- anova_inducedt0[[1]][["Pr(>F)"]][1]

tukey_inducedt0 <- HSD.test(aov(mg_co_induced ~ treatment +rep, data=respirationt0), "treatment", alpha = 0.05)$groups %>% rownames_to_column(var= "treatment")
tukey_inducedt0

meaninducedt0 <- aggregate(mg_co_induced ~ treatment, data=respirationt0, FUN= mean)

plot_induced_t0 <- respirationt0 %>% ggplot(aes(x= factor(treatment, level= treat.order), y= mg_co_induced, fill=treatment))+
  geom_point(pch= 24, size= 7)+ 
  scale_fill_manual(values= treat.colors)+
  xlab("")+
  ylab("mg C-CO2/kg dry soil")+#induced
  stat_summary(fun=mean, aes(x= treatment, y = mg_co_induced),
               geom= "crossbar", colour= "black", width= 0.7, data=meaninducedt0)+
  geom_text(data=tukey_inducedt0, aes(x=treatment, y = mg_co_induced, label= groups), 
            size= 4, colour="black", fontface = "bold", nudge_x= 0.2, nudge_y= 4.5)+
  labs(subtitle = "" , caption = "Different letters indicate statistically significant differences between each treatment at T0.\nCrossbar indicate the average of the three samples ")+
  theme(plot.title = element_text(size=1))+
  ggtitle("Induced respiration at T0", subtitle= paste("ANOVA (HSDTukey) p-value: ", format(pvalue_anova_inducedt0, scientific= TRUE, digits=2 ), sep= ""))+
  scale_x_discrete(labels= treat.labels)+
  guides(fill=guide_legend(title="Treatment"))+
  theme_classic2()
plot_induced_t0
ggsave(plot_induced_t0, file="soil_respiration_result/dif_induced_t0.png", width= 1900,height= 1200, units= "px")

```


### Differences of respiration between treatments at T1


```{r respiration basal and induced t1, echo=FALSE}

respirationt1 <- subset(soil_resp, soil_resp$time == "t1")
#analized basal
shapiro.test(respirationt1$mg_co_basal)#data basal normal
leveneTest(mg_co_basal~ treatment, data=respirationt0) #equal Variance 

anova_basalt1 <- aov(mg_co_basal ~ treatment + rep, data=respirationt1) #signification different 
anova_basalt1 <- summary(aov(mg_co_basal ~ treatment + rep, data=respirationt1))

pvalue_anova_basalt1 <- anova_basalt1[[1]][["Pr(>F)"]][1]
tukey_basalt1 <- HSD.test(aov(mg_co_basal ~ treatment + rep, data=respirationt1), "treatment", alpha = 0.05)$groups %>% rownames_to_column(var= "treatment")
tukey_basalt1

meanbasalt1 <- aggregate(mg_co_basal ~ treatment, data=respirationt1, FUN= mean)

plot_basal_t1 <- respirationt1 %>% ggplot(aes(x= factor(treatment, level= treat.order), y= mg_co_basal, fill=treatment))+
  geom_point(pch= 24, size= 7)+ 
  scale_fill_manual(values= treat.colors)+
  guides(fill=guide_legend(title="Treatment"))+
  xlab("")+
  ylab("mg C-CO2/kg dry soil")+#basal
  stat_summary(fun=mean, aes(x= treatment, y = mg_co_basal),
               geom= "crossbar", colour= "black", width= 0.7, data=meanbasalt1)+
  geom_text(data=tukey_basalt1, aes(x=treatment, y = mg_co_basal, label= groups), 
            size= 4, colour="black", fontface = "bold", nudge_x= 0.2, nudge_y= 0.6)+
  labs(subtitle = "" , caption = "Different letters indicate statistically significant differences between each treatment at T1.\nCrossbar indicate the average of the three samples ")+
  theme(plot.title = element_text(size=1))+
  scale_x_discrete(labels= treat.labels)+
  ggtitle("Basal respiration at T1", subtitle= paste("ANOVA (HSDTukey) p-value: ", format(pvalue_anova_basalt1, scientific= TRUE, digits=2 ), sep= ""))+
  theme_classic2()
plot_basal_t1
ggsave(plot_basal_t1, file="soil_respiration_result/dif_basal_t1.png", width= 1900,height= 1200, units= "px")

#analized induced
shapiro.test(respirationt1$mg_co_induced)#data induced no normal
soil_resp %>% group_by(treatment) %>% shapiro_test(mg_co_induced) #all data of each time of induced are normal
leveneTest(mg_co_induced~ treatment, data=respirationt1) #equal variance
anova_inducedt1 <- aov(mg_co_induced ~ treatment +rep, data=respirationt1)#signification different 
anova_inducedt1 <-summary(aov(mg_co_induced ~ treatment +rep, data=respirationt1))


pvalue_anova_inducedt1 <- anova_inducedt1[[1]][["Pr(>F)"]][1]

tukey_inducedt1 <- HSD.test(aov(mg_co_induced ~ treatment +rep, data=respirationt1), "treatment", alpha = 0.05)$groups %>% rownames_to_column(var= "treatment")
tukey_inducedt1

meaninducedt1 <- aggregate(mg_co_induced ~ treatment, data=respirationt1, FUN= mean)

plot_induced_t1 <- respirationt1 %>% ggplot(aes(x= factor(treatment, level= treat.order), y= mg_co_induced, fill=treatment))+
  geom_point(pch= 24, size= 7)+ 
  scale_fill_manual(values= treat.colors)+
  guides(fill=guide_legend(title="Treatment"))+
  xlab("")+
  ylab("mg C-CO2/kg dry soil")+#induced
  stat_summary(fun=mean, aes(x= treatment, y = mg_co_induced),
               geom= "crossbar", colour= "black", width= 0.7, data=meaninducedt1)+
  geom_text(data=tukey_inducedt1, aes(x=treatment, y = mg_co_induced, label= groups), 
            size= 4, colour="black", fontface = "bold", nudge_x= 0.2, nudge_y= 5)+
  labs(subtitle = "" , caption = "Different letters indicate statistically significant differences between each treatment at T1.\nCrossbar indicate the average of the three samples ")+
  ggtitle("Induced respiration at T1", subtitle= paste("ANOVA (HSDTukey) p-value: ", format(pvalue_anova_inducedt1, scientific= TRUE, digits=2 ), sep= ""))+
  theme(plot.title = element_text(size=1))+
  scale_x_discrete(labels= treat.labels)+
  theme_classic2()
plot_induced_t1

ggsave(plot_induced_t1, file="soil_respiration_result/dif_induced_t1.png", width= 1900,height= 1200, units= "px")

```

### Differences of respiration between treatments at T2


```{r respiration in t2, echo=FALSE}

respirationt2 <- subset(soil_resp, soil_resp$time == "t2")
#analized basal
shapiro.test(respirationt2$mg_co_basal)#data basal normal
leveneTest(mg_co_basal~ treatment, data=respirationt0) #equal Variance 

anova_basalt2 <- aov(mg_co_basal ~ treatment + rep, data=respirationt2) #signification different 
anova_basalt2 <- summary(aov(mg_co_basal ~ treatment + rep, data=respirationt2))
pvalue_anova_basalt2 <- anova_basalt2[[1]][["Pr(>F)"]][1]
tukey_basalt2 <- HSD.test(aov(mg_co_basal ~ treatment + rep, data=respirationt2), "treatment", alpha = 0.05)$groups %>% rownames_to_column(var= "treatment")
tukey_basalt2

meanbasalt2 <- aggregate(mg_co_basal ~ treatment, data=respirationt2, FUN= mean)

plot_basal_t2 <- respirationt2 %>% ggplot(aes(x= factor(treatment, level= treat.order), y= mg_co_basal, fill=treatment))+
  geom_point(pch= 24, size= 7)+ 
  scale_fill_manual(values= treat.colors)+
  guides(fill=guide_legend(title="Treatment"))+
  xlab("")+
  ylab("mg C-CO2/kg dry soil")+#basal
  stat_summary(fun=mean, aes(x= treatment, y = mg_co_basal),
               geom= "crossbar", colour= "black", width= 0.7, data=meanbasalt2)+
  geom_text(data=tukey_basalt2, aes(x=treatment, y = mg_co_basal, label= groups), 
            size= 4, colour="black", fontface = "bold", nudge_x= 0.2, nudge_y= 1)+
  theme(plot.title = element_text(size=1))+
  scale_x_discrete(labels= treat.labels)+
  labs(subtitle = "" , caption = "Different letters indicate statistically significant differences between each treatment at T2.\nCrossbar indicate the average of the three samples ")+
  ggtitle("Basal respiration at T2", subtitle= paste("ANOVA (HSDTukey) p-value: ", format(pvalue_anova_basalt2, scientific= TRUE, digits=2 ), sep= ""))+
  theme_classic2()
plot_basal_t2

ggsave(plot_basal_t2, file="soil_respiration_result/dif_basal_t2.png", width= 1900,height= 1200, units= "px")

#analized induced
shapiro.test(respirationt2$mg_co_induced)#data induced no normal
soil_resp %>% group_by(treatment) %>% shapiro_test(mg_co_induced) #all data of each time of induced are normal
leveneTest(mg_co_induced~ treatment, data=respirationt2) #equal variance
anova_inducedt2 <- aov(mg_co_induced ~ treatment +rep, data=respirationt2)#signification different 
anova_inducedt2<-summary(aov(mg_co_induced ~ treatment +rep, data=respirationt2))
pvalue_anova_inducedt2 <- anova_inducedt2[[1]][["Pr(>F)"]][1]
tukey_inducedt2 <- HSD.test(aov(mg_co_induced ~ treatment +rep, data=respirationt2), "treatment", alpha = 0.05)$groups %>% rownames_to_column(var= "treatment")
tukey_inducedt2

meaninducedt2 <- aggregate(mg_co_induced ~ treatment, data=respirationt2, FUN= mean)

plot_induced_t2 <- respirationt2 %>% ggplot(aes(x= factor(treatment, level= treat.order), y= mg_co_induced, fill=treatment))+
  geom_point(pch= 24, size= 7)+ 
  scale_fill_manual(values= treat.colors)+
  guides(fill=guide_legend(title="Treatment"))+
  xlab("")+
  ylab("mg C-CO2/kg dry soil")+#induced
  stat_summary(fun=mean, aes(x= treatment, y = mg_co_induced),
               geom= "crossbar", colour= "black", width= 0.7, data=meaninducedt2)+
  geom_text(data=tukey_inducedt2, aes(x=treatment, y = mg_co_induced, label= groups), 
            size= 4, colour="black", fontface = "bold", nudge_x= 0.2, nudge_y= 4.5)+
  theme(plot.title = element_text(size=1))+
  scale_x_discrete(labels= treat.labels)+
  labs(subtitle = "" , caption = "Different letters indicate statistically significant differences between each treatment at T2.\nCrossbar indicate the average of the three samples ")+
  ggtitle("Induced respiration at T2", subtitle= paste("ANOVA (HSDTukey) p-value: ", format(pvalue_anova_inducedt2, scientific= TRUE, digits=2 ), sep= ""))+
  theme_classic2()
plot_induced_t2
ggsave(plot_induced_t2, file="soil_respiration_result/dif_induced_t2.png", width= 1900,height= 1200, units= "px")

```

### Differences of respiration between treatments at T3

```{r respiration at t3, echo=FALSE}
respirationt3 <- subset(soil_resp, soil_resp$time == "t3")
#analized basal
shapiro.test(respirationt3$mg_co_basal)#data basal normal
leveneTest(mg_co_basal~ treatment, data=respirationt0) #equal Variance 

anova_basalt3 <- aov(mg_co_basal ~ treatment +rep, data=respirationt3) #signification different 
anova_basalt3<- summary(aov(mg_co_basal ~ treatment +rep, data=respirationt3))
pvalue_anova_basalt3 <- anova_basalt3[[1]][["Pr(>F)"]][1]
tukey_basalt3 <- HSD.test(aov(mg_co_basal ~ treatment +rep, data=respirationt3), "treatment", alpha = 0.05)$groups %>% rownames_to_column(var= "treatment")
tukey_basalt3

meanbasalt3 <- aggregate(mg_co_basal ~ treatment, data=respirationt3, FUN= mean)

plot_basal_t3 <- respirationt3 %>% ggplot(aes(x= factor(treatment, level= treat.order), y= mg_co_basal, fill=treatment))+
  geom_point(pch= 24, size= 7)+ 
  scale_fill_manual(values= treat.colors)+
  guides(fill=guide_legend(title="Treatment"))+
  xlab("")+
  ylab("mg C-CO2/kg dry soil")+#basal
  stat_summary(fun=mean, aes(x= treatment, y = mg_co_basal),
               geom= "crossbar", colour= "black", width= 0.7, data=meanbasalt3)+
  geom_text(data=tukey_basalt3, aes(x=treatment, y = mg_co_basal, label= groups), 
            size= 4, colour="black", fontface = "bold", nudge_x= 0.2, nudge_y= 1.5)+
  theme(plot.title = element_text(size=1))+
  scale_x_discrete(labels= treat.labels)+
  labs(subtitle = "" , caption = "Different letters indicate statistically significant differences between each treatment at T3.\nCrossbar indicate the average of the three samples ")+
  ggtitle("Basal respiration at T3", subtitle= paste("ANOVA (HSDTukey) p-value: ", format(pvalue_anova_basalt3, scientific= TRUE, digits=2 ), sep= ""))+
  theme_classic2()

plot_basal_t3
ggsave(plot_basal_t3, file="soil_respiration_result/dif_basal_t3.png", width= 1900,height= 1200, units= "px")
#analized induced
shapiro.test(respirationt3$mg_co_induced)#data induced no normal
soil_resp %>% group_by(treatment) %>% shapiro_test(mg_co_induced) #all data of each time of induced are normal
leveneTest(mg_co_induced~ treatment, data=respirationt3) #equal variance
anova_inducedt3 <- aov(mg_co_induced ~ treatment +rep, data=respirationt3)#signification different 
anova_inducedt3<- summary(aov(mg_co_induced ~ treatment +rep, data=respirationt3))
pvalue_anova_inducedt3 <- anova_inducedt3[[1]][["Pr(>F)"]][1]
tukey_inducedt3 <- HSD.test(aov(mg_co_induced ~ treatment +rep, data=respirationt3), "treatment", alpha = 0.05)$groups %>% rownames_to_column(var= "treatment")
tukey_inducedt3

meaninducedt3 <- aggregate(mg_co_induced ~ treatment, data=respirationt3, FUN= mean)

plot_induced_t3 <- respirationt3 %>% ggplot(aes(x= factor(treatment, level= treat.order), y= mg_co_induced, fill=treatment))+
  geom_point(pch= 24, size= 7)+ 
  scale_fill_manual(values= treat.colors)+
  guides(fill=guide_legend(title="Treatment"))+
  xlab("")+
  ylab("mg C-CO2/kg dry soil")+#induced
  stat_summary(fun=mean, aes(x= treatment, y = mg_co_induced),
               geom= "crossbar", colour= "black", width= 0.7, data=meaninducedt3)+
  geom_text(data=tukey_inducedt3, aes(x=treatment, y = mg_co_induced, label= groups), 
            size= 4, colour="black", fontface = "bold", nudge_x= 0.2, nudge_y= 4.5)+
  theme(plot.title = element_text(size=1))+
  scale_x_discrete(labels= treat.labels)+
  labs(subtitle = "" , caption = "Different letters indicate statistically significant differences between each treatment at T3.\nCrossbar indicate the average of the three samples ")+
  ggtitle("Induced respiration at T3", subtitle= paste("ANOVA (HSDTukey) p-value: ", format(pvalue_anova_inducedt3, scientific= TRUE, digits=2 ), sep= ""))+
  theme_classic2()

plot_induced_t3
ggsave(plot_induced_t3, file="soil_respiration_result/dif_induced_t3.png", width= 1900,height= 1200, units= "px")

```


### Differences of respiration between treatments at T4


```{r respiration at t4, echo=FALSE}
respirationt4 <- subset(soil_resp, soil_resp$time == "t4")
#analized basal
shapiro.test(respirationt4$mg_co_basal)#data basal normal
leveneTest(mg_co_basal~ treatment, data=respirationt0) #equal Variance 

anova_basalt4 <- aov(mg_co_basal ~ treatment +rep, data=respirationt4) #signification different 
anova_basalt4<-summary(aov(mg_co_basal ~ treatment +rep, data=respirationt4))
pvalue_anova_basalt4 <- anova_basalt4[[1]][["Pr(>F)"]][1]
tukey_basalt4 <- HSD.test(aov(mg_co_basal ~ treatment +rep, data=respirationt4), "treatment", alpha = 0.05)$groups %>% rownames_to_column(var= "treatment")
tukey_basalt4

meanbasalt4 <- aggregate(mg_co_basal ~ treatment, data=respirationt4, FUN= mean)

plot_basal_t4 <- respirationt4 %>% ggplot(aes(x= factor(treatment, level= treat.order), y= mg_co_basal, fill=treatment))+
  geom_point(pch= 24, size= 7)+ 
  scale_fill_manual(values= treat.colors)+
  guides(fill=guide_legend(title="Treatment"))+
  xlab("")+
  ylab("mg C-CO2/kg dry soil")+#basal
  stat_summary(fun=mean, aes(x= treatment, y = mg_co_basal),
               geom= "crossbar", colour= "black", width= 0.7, data=meanbasalt4)+
  geom_text(data=tukey_basalt4, aes(x=treatment, y = mg_co_basal, label= groups), 
            size= 4, colour="black", fontface = "bold", nudge_x= 0.2, nudge_y= 1)+
  theme(plot.title = element_text(size=1))+
  scale_x_discrete(labels= treat.labels)+
  labs(subtitle = "" , caption = "Different letters indicate statistically significant differences between each treatment at T4.\nCrossbar indicate the average of the three samples ")+
  ggtitle("Basal respiration at T4", subtitle= paste("ANOVA (HSDTukey) p-value: ", format(pvalue_anova_basalt4, scientific= TRUE, digits=2 ), sep= ""))+
  theme_classic2()
plot_basal_t4
ggsave(plot_basal_t4, file="soil_respiration_result/dif_basal_t4.png", width= 1900,height= 1200, units= "px")

#analized induced
shapiro.test(respirationt4$mg_co_induced)#data induced no normal
soil_resp %>% group_by(treatment) %>% shapiro_test(mg_co_induced) #all data of each time of induced are normal
leveneTest(mg_co_induced~ treatment, data=respirationt4) #equal variance
anova_inducedt4 <- aov(mg_co_induced ~ treatment + rep, data=respirationt4)#signification different 
anova_inducedt4 <- summary(aov(mg_co_induced ~ treatment + rep, data=respirationt4))
pvalue_anova_inducedt4 <- anova_inducedt4[[1]][["Pr(>F)"]][1]
tukey_inducedt4 <- HSD.test(aov(mg_co_induced ~ treatment + rep, data=respirationt4), "treatment", alpha = 0.05)$groups %>% rownames_to_column(var= "treatment")
tukey_inducedt4

meaninducedt4 <- aggregate(mg_co_induced ~ treatment, data=respirationt4, FUN= mean)

plot_induced_t4 <- respirationt4 %>% ggplot(aes(x= factor(treatment, level= treat.order), y= mg_co_induced, fill=treatment))+
  geom_point(pch= 24, size= 7)+ 
  scale_fill_manual(values= treat.colors)+
  guides(fill=guide_legend(title="Treatment"))+
  xlab("")+
  ylab("mg C-CO2/kg dry soil")+#induced
  stat_summary(fun=mean, aes(x= treatment, y = mg_co_induced),
               geom= "crossbar", colour= "black", width= 0.7, data=meaninducedt4)+
  geom_text(data=tukey_inducedt4, aes(x=treatment, y = mg_co_induced, label= groups), 
            size= 4, colour="black", fontface = "bold", nudge_x= 0.2, nudge_y= 4.5)+
  theme(plot.title = element_text(size=1))+
  scale_x_discrete(labels= treat.labels)+
  labs(subtitle = "" , caption = "Different letters indicate statistically significant differences between each treatment at T4.\nCrossbar indicate the average of the three samples ")+
  ggtitle("Induced respiration at T4", subtitle= paste("ANOVA (HSDTukey) p-value: ", format(pvalue_anova_inducedt4, scientific= TRUE, digits=2 ), sep= ""))+
  theme_classic2()
plot_induced_t4
ggsave(plot_induced_t4, file="soil_respiration_result/dif_induced_t4.png", width= 1900,height= 1200, units= "px")

```
¨

```{r graph basal, echo=FALSE}

respiration_basal_all <- ggarrange(plot_basal_t0 + theme(legend.position="none"),  plot_basal_t1 + theme(legend.position="none"), plot_basal_t2 + theme(legend.position="none"),plot_basal_t3 + theme(legend.position="none"), plot_basal_t4 + theme(legend.position="none"))
respiration_basal_all                                  

ggsave(respiration_basal_all, file="soil_respiration_result/all_basal_respiration.png", width= 3800,height= 3600, units= "px")


respiration_induced_all <- ggarrange(plot_induced_t0 + theme(legend.position="none"), plot_induced_t1 + theme(legend.position="none"), plot_induced_t2 + theme(legend.position="none"), plot_induced_t3+ theme(legend.position="none"), plot_induced_t4 + theme(legend.position="none"))
respiration_induced_all
ggsave(respiration_induced_all, file="soil_respiration_result/all_induced_respiration.png", width= 1900,height= 1200, units= "px")

```



### How much has the respiration of each treatment varied from the beginning of the experiment to the end (T4 and tT0)?

Delta Variation (ΔBasal Respiration): It represents the change or difference in basal respiration values between two specific time points. Mathematically, it is often expressed as the subtraction of the basal respiration value at an earlier time point from the value at a later time point.

-Positive ΔBasal Respiration: Indicates an increase in basal respiration from the earlier time point to the later time point. This could be due to factors such as increased microbial activity, higher availability of organic matter, or changes in environmental conditions that favor microbial respiration.

-Negative ΔBasal Respiration: Indicates a decrease in basal respiration from the earlier time point to the later time point. This could be due to factors such as decreased microbial activity, depletion of organic matter, or changes in environmental conditions that suppress microbial respiration.

-Zero ΔBasal Respiration: Indicates no change in basal respiration between the two time points. The microbial activity and respiration rates remain constant over this period.

```{r change between t0 and t4, echo=FALSE}

treat.pot <- paste(soil_resp$treatment, soil_resp$rep, sep= "_") #para mantener el error
soil_resp <- data.frame(soil_resp, treat.pot)
respirationt0t4 <- soil_resp %>% filter(time == "t0" | time == "t4") #data only t0 and t4
respirationt0t4mean <- soil_resp_mean %>% filter(time == "t0" | time == "t4") #data only t0 and t4
#basal
shapiro.test(respirationt0t4$mg_co_basal)#normal data
leveneTest(mg_co_basal ~ time * treatment, data=respirationt0t4) #equal variances
#modelo 
model_respibasalt0t4 <- lmer(mg_co_basal~ time*treatment + (1|treat.pot), data=respirationt0t4)
summary(model_respibasalt0t4)

anova_model_respibasalt0t4 <- anova(model_respibasalt0t4)
anova_model_respibasalt0t4 #p value 0.000939
emm_respibasalt0t4 <- emmeans(model_respibasalt0t4, pairwise ~ time | treatment)
summary(emm_respibasalt0t4)


emm_respibasalt0t4$contrasts %>% rbind(adjust= "fdr") %>% summary(infer = TRUE)#el adjust es para el metodo de los pvalues
basalt0t4contraste <- data.frame(emm_respibasalt0t4$contrasts)#data frame para sacar el valor del contraste 
basalt0t4contraste
basalt0t4emmeans <- data.frame(emm_respibasalt0t4$emmeans) %>% group_by(treatment) %>% summarise(across(.cols= "emmean",.fns=mean),.groups = "drop")
basalt0t4emmeans

emm_respibasalt0t4 <- cld(emmeans(model_respibasalt0t4, specs = pairwise~treatment), Letters= letters, reversed= TRUE) #add letters
emm_respibasalt0t4
pval_basalt0t4_time <- data.frame(treatment= basalt0t4contraste$treatment, pval= basalt0t4contraste$p.value, emmean= basalt0t4emmeans$emmean)
pval_basalt0t4_time$sign <- ifelse(pval_basalt0t4_time$pval>0.05, paste("ns"),
                                 ifelse(pval_basalt0t4_time$pval>0.01, paste("*"),
                                        ifelse(pval_basalt0t4_time$pval> 0.001,
                                               paste("**"), paste("***"))))
basalt0t4_time <- respirationt0t4mean %>% ggplot(aes(x=factor(treatment, level= treat.order), y= mg_co_basal_mean))+
  geom_col(aes(fill= time), col= "black", position= "dodge")+
  geom_text(data = pval_basalt0t4_time, aes(x= treatment, y= emmean, label= sign), nudge_x= 0, nudge_y= 10)+
  xlab("")+
  ylab("basal mg C-CO2/kg dry soil")+
  theme_classic2()+
  scale_x_discrete(labels= treat.labels)+
  scale_fill_manual(values = c("#A1D5D2", "#8C8380"))+
  labs(subtitle = "" , caption = "Different (*) indicates differences between the levels of the treatment factor at each time (T0-T4).\n caculated by Estimated Marginal Means posthoc -Tukey's adjustment method") +
  ggtitle("Basal respiration in T0 and T4\nMean and SD", subtitle= paste("LMEs-Anova III-Time:Treatment(Tukey) p-value: ", format(anova_model_respibasalt0t4$`Pr(>F)`[[3]], scientific= TRUE, digits=2 ), sep= ""))+
  geom_errorbar(data=respirationt0t4mean, aes(ymin=mg_co_basal_mean-mg_co_basal_sd, 
            ymax=mg_co_basal_mean+mg_co_basal_sd, group= time),position=position_dodge(0.9),
                                            size =0.2, width=0.2)+
  guides(fill=guide_legend(title="Time"))+
  theme(
    plot.title = element_text(size = 12),
    plot.subtitle = element_text(size = 11),
    plot.caption = element_text(size = 9))

basalt0t4_time
ggsave(basalt0t4_time, file="soil_parameters_result/basalt0t4_time.png", width= 1900,height= 1200, units= "px")
#delta
delta_basalt0t4 <- data.frame (treatment=respirationt0$treatment, t4= respirationt4$mg_co_basal, t0=respirationt0$mg_co_basal)
delta_basalt0t4$delta <- delta_basalt0t4$t4-delta_basalt0t4$t0
delta_basalt0t4

delta_basalt0t4_modelo <- lm(delta~ treatment, data=delta_basalt0t4)#model
anova_delta_basalt0t4_modelo<-anova(delta_basalt0t4_modelo)

emm.delta.basalt0t4 <- emmeans(delta_basalt0t4_modelo, pairwise ~  treatment)
emm.delta.basalt0t4
emm.delta.basalt0t4$contrasts %>% rbind(adjust= "fdr") %>% 
  summary(infer = TRUE)
emm.delta.basalt0t4 <- cld(emmeans(delta_basalt0t4_modelo, specs = pairwise~treatment), Letters= letters, reversed= TRUE)

plotdelta_basalt0t4 <- delta_basalt0t4 %>% ggplot(aes(x=factor(treatment, level= treat.order), y= delta, fill= treatment))+
  geom_point(pch= 24, size= 7)+
  xlab("")+
  ylab("Delta variation between t4-t0")+
  geom_hline(yintercept= 0, linetype = "dashed", color = "red")+
  scale_fill_manual(values= treat.colors)+
  geom_text(data= emm.delta.basalt0t4, aes(x= treatment, y= emmean, label= `.group`), nudge_x = 0.2, nudge_y = 0.2)+
  scale_x_discrete(labels= treat.labels)+
  labs(subtitle = "" , caption = "Different letters indicates differences between the levels of the treatment factor in delta variation (T4-T0).\n caculated by Estimated Marginal Means post hoc -Tukey's adjustment method.")+
  ggtitle("Δ Basal Respiration at T0 and T4", subtitle= paste("LinealModel-Anova Treatment(Tukey) p-value: ", format(anova_delta_basalt0t4_modelo$`Pr(>F)`[[1]], scientific= TRUE, digits=2 ), sep= ""))+
  theme_classic2()+
  guides(fill=guide_legend(title="Treatment"))
plotdelta_basalt0t4

basal_t0t4plot <- ggarrange(basalt0t4_time, plotdelta_basalt0t4+ guides(fill= "none"), ncol = 2)
basal_t0t4plot
#basal_t0t4plot <- annotate_figure(basal_t0t4plot, top = text_grob("Basal respiration", color = "black", face = "bold", size = 10),bottom = text_grob("Interaction and variation between beginning and end of the experiment", size=10)) para dar titulo

ggsave(basal_t0t4plot, file="soil_respiration_result/interac_basal_t4.png", width= 3800,height= 2200, units= "px")

#induced

shapiro.test(respirationt0t4$mg_co_induced)# no normal data, but all data of each time of induced are normal
leveneTest(mg_co_induced ~ time * treatment, data=respirationt0t4) #equal variances
#modelo 
model_respi_inducet0t4 <- lmer(mg_co_induced~ time*treatment + (1|treat.pot), data=respirationt0t4)
summary(model_respi_inducet0t4)

anova_model_respi_inducet0t4 <- anova(model_respi_inducet0t4)
anova_model_respi_inducet0t4 #p value 0.0001959 time:treatment
emm_respi_inducet0t4 <- emmeans(model_respi_inducet0t4, pairwise ~ time | treatment)
summary(emm_respi_inducet0t4)


emm_respi_inducet0t4$contrasts %>% rbind(adjust= "fdr") %>% summary(infer = TRUE)#el adjust es para el metodo de los pvalues
inducet0t4contraste <- data.frame(emm_respi_inducet0t4$contrasts)#data frame para sacar el valor del contraste 
inducet0t4contraste
inducet0t4emmeans <- data.frame(emm_respi_inducet0t4$emmeans) %>% group_by(treatment) %>% summarise(across(.cols= "emmean",.fns=mean),.groups = "drop")
inducet0t4emmeans

emm_respi_inducet0t4 <- cld(emmeans(model_respi_inducet0t4, specs = pairwise~treatment), Letters= letters, reversed= TRUE) #add letters
emm_respi_inducet0t4
pval_inducet0t4_time <- data.frame(treatment= inducet0t4contraste$treatment, pval= inducet0t4contraste$p.value, emmean= inducet0t4emmeans$emmean)
pval_inducet0t4_time$sign <- ifelse(pval_inducet0t4_time$pval>0.05, paste("ns"),
                                 ifelse(pval_inducet0t4_time$pval>0.01, paste("*"),
                                        ifelse(pval_inducet0t4_time$pval> 0.001,
                                               paste("**"), paste("***"))))
inducet0t4_time <- respirationt0t4mean %>% ggplot(aes(x=factor(treatment, level= treat.order), y= mg_co_induced_mean))+
  geom_col(aes(fill= time), col= "black", position= "dodge")+
  geom_text(data = pval_inducet0t4_time, aes(x= treatment, y= emmean, label= sign), nudge_x= 0, nudge_y= 10)+
  xlab("")+
  ylab("mg C-CO2/kg dry soil")+
  theme_classic2()+
  scale_x_discrete(labels= treat.labels)+
  scale_fill_manual(values = c("#A1D5D2", "#8C8380"))+
  labs(subtitle = "" , caption = "Different (*) indicates differences between the levels of the treatment factor at each time (T0-T4).\n caculated by Estimated Marginal Means posthoc -Tukey's adjustment method") +
  ggtitle("Induced respiration in T0 and T4\nMean and SD", subtitle= paste("LMEs-Anova III-Time:Treatment(Tukey) p-value: ", format(anova_model_respi_inducet0t4$`Pr(>F)`[[3]], scientific= TRUE, digits=2 ), sep= ""))+
  geom_errorbar(data=respirationt0t4mean, aes(ymin=mg_co_induced_mean-mg_co_induced_sd, 
            ymax=mg_co_induced_mean+mg_co_induced_sd, group= time),position=position_dodge(0.9),
                                            size =0.2, width=0.2)+
  guides(fill=guide_legend(title="Time"))+
  theme(
    plot.title = element_text(size = 12),
    plot.subtitle = element_text(size = 11),
    plot.caption = element_text(size = 9))

inducet0t4_time
ggsave(inducet0t4_time, file="soil_parameters_result/inducet0t4_time.png", width= 1900,height= 1200, units= "px")
#delta
delta_inducet0t4 <- data.frame (treatment=respirationt0$treatment, t4= respirationt4$mg_co_induced, t0=respirationt0$mg_co_induced)
delta_inducet0t4$delta <- delta_inducet0t4$t4-delta_inducet0t4$t0
delta_inducet0t4

delta_inducet0t4_modelo <- lm(delta~ treatment, data=delta_inducet0t4)#model
anova_delta_inducet0t4_modelo<-anova(delta_inducet0t4_modelo)

emm.delta.inducet0t4 <- emmeans(delta_inducet0t4_modelo, pairwise ~  treatment)
emm.delta.inducet0t4
emm.delta.inducet0t4$contrasts %>% rbind(adjust= "fdr") %>% 
  summary(infer = TRUE)
emm.delta._inducet0t4 <- cld(emmeans(delta_inducet0t4_modelo, specs = pairwise~treatment), Letters= letters, reversed= TRUE)

plotdelta_inducet0t4 <- delta_inducet0t4 %>% ggplot(aes(x=factor(treatment, level= treat.order), y= delta, fill= treatment))+
  geom_point(pch= 24, size= 7)+
  xlab("")+
  ylab("Delta variation between t4-t0")+
  geom_hline(yintercept= 0, linetype = "dashed", color = "red")+
  scale_fill_manual(values= treat.colors)+
  geom_text(data= emm.delta._inducet0t4, aes(x= treatment, y= emmean, label= `.group`), nudge_x = 0.2, nudge_y = 0.2)+
  scale_x_discrete(labels= treat.labels)+
  labs(subtitle = "" , caption = "Different letters indicates differences between the levels of the treatment factor in delta variation (T4-T0).\n caculated by Estimated Marginal Means post hoc -Tukey's adjustment method.")+
  ggtitle("Δ Induced Respiration at T0 and T4", subtitle= paste("LinealModel-Anova Treatment(Tukey) p-value: ", format(anova_delta_inducet0t4_modelo$`Pr(>F)`[[1]], scientific= TRUE, digits=2 ), sep= ""))+
  theme_classic2()+
  guides(fill=guide_legend(title="Treatment"))
plotdelta_inducet0t4

induce_t0t4plot <- ggarrange(inducet0t4_time, plotdelta_inducet0t4+ guides(fill= "none"), ncol = 2)
induce_t0t4plot
#_induce_t0t4plot <- annotate_figure(_induce_t0t4plot, top = text_grob("_induce respiration", color = "black", face = "bold", size = 10),bottom = text_grob("Interaction and variation between beginning and end of the experiment", size=10)) para dar titulo

ggsave(induce_t0t4plot, file="soil_respiration_result/interac_inducet0t4.png", width= 3800,height= 2200, units= "px")


```

